Appendix of Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-11-17 manuscript/bef_stability/appendixB.Rmd
Update appendix

---
title: Appendix
author: Alain Danet, Maud Mouchet, Elisa Thebault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))
source(mypath("R", "cleaning_methods.R"))

@@

@@

@@

Diet information were obtained from fishbase.org and literature [REF willem?]. The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had as strict piscivorous
life stage. Fish species were largely omnivorous.

@@

```{r, fig.height = 20/2.54}
pca_pressure <- habitat_press
myload(fourier, dir = mypath("data-raw", "polluants"))
myload(flow_temp_metrics, station_analysis, the_8_hydrologic_basin, dir = mypath("data"))

@@

map_pressure <- map(var_to_plot,
~p +
  geom_sf(data = sf_pressure, aes_string(color = .x), size = .5) +
  viridis::scale_color_viridis() +
  labs(title = str_replace_all(.x, var_replacement)) +
  theme(legend.title = element_blank(),
    text = element_text(size = 8),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
 guides(fill = guide_colourbar(barwidth = 0.5, barheight = 10))
)

@@

We reconstructed the stream network and the connection between sites with the
`openSTARS` R package. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(IGN)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network [@pella_reseau_2012].

@@

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their prediction performance. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014, @hoef_ssn_2014].

@@

# Species list and food diet data

```{r sp-list}
myload(species, fish_diet_shift, dir = mypath("data"))
kable(species,
  format = "latex",
  booktabs = T,
  label = paste0("fish-sp"),
  caption = "Fish species list."
  ) %>%
kable_styling(latex_options =c("striped"))
```

```{r food}
kable(select(fish_diet_shift, - species_name, -remark),
  format = "latex",
  booktabs = T,
  label = paste0("onto-diet"),
  caption = "Ontogenic diet shift for each species"
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

# Analysis

## Time series classication

@@

## PCA of environmental variables

```{r eig-plot}
rotated_eigen_values <- 
  tibble(
    name = paste0("RC", seq(1,14)),
    type = c(rep("Kept", 2), rep("No kept", 14 - 2)),
    value = pca_rotated$rotated$values
  )
eigen_rotated_plot <-
  ggplot(rotated_eigen_values,
    aes(x = reorder(name, -value) , y = value)) +
  geom_col(aes(fill = type)) +
  scale_fill_discrete(name = "Status") +
  labs(x = "Rotated component", y = "Eigen values")
```

@@

```{r}
corr_pca_caption <- paste0(
  "Spearman correlation between environmental variables used in the PCA.
  The variables have been scaled before to compute the correlation."
)

@@

```{r rc2-map, fig.cap = caption_rc2_map}
p +
  geom_sf(data = rc2_sf, aes(color = log(value + abs(min(value)) + 1 ))) +
  viridis::scale_color_viridis() +
  labs(title = "RC2 (Log)") +
  theme(legend.title = element_blank())
```

@@

## Model summary and diagnostics

### Data transformation

Additionally, we logged the axis one, two and three to normalize their
distribution and we reversed the third axis to get increasing average
temperature with increasing values of the axis and so decreasing values of
altitude.

```{r}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 5)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 5), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c(rep("Log base 10", 3), rep("No", 2), "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  interpretation = c(rep("", 11)) 
) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "Variable description and transformation"
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
sem_variable_table
```

@@

---
# Species richness to community attributes 
---
To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. In particular, species richness decreased connectance
following results from @thebault_stability_2010; @winemiller_must_1989. But
theoretically, some ecological networks can show the inverse relationship
[@winemiller_must_1989]. (6) We hypothesized also that species richness will
modulate average weighted trophic level because of food web assembly processes.
A community with more fish species can support higher trophic level. Finally, we
set up causal relationships from community attributes to stability components.

@@

## Model linear Relationship

### Variables

@@

### Colinearity tables

@@

### Biomass and stability models

@@

kable(format_coef_table(stab_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_stab_sem_caption
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

@@

kable(format_coef_table(bm_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_bm_sem_caption
  ) %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r rsq-model}
rm_n <- str_remove_all(get_sem_var_name_replacement(), pattern = "\n")
names(rm_n) <- names(get_sem_var_name_replacement()) 
rsq_table <- map_dfr(mod_stab_list, piecewiseSEM::rsquared, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2)) %>%
  mutate(Response = str_replace_all(Response, rm_n)) %>%
  dplyr::select(Response, Marginal, Conditional, R.squared)
rsq_table_caption <- paste0(
  "R squared of the linear models used in the structural equation modeling."
)
kable(format_coef_table(bm_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_bm_sem_caption
  ) %>%
kable_styling(latex_options = c("striped"))
```

@@

get_intercept_variance <- function (mod) {
  vc <- VarCorr(mod)
  suppressWarnings(storage.mode(vc) <- "numeric")
  enframe(vc[1,])
}
random_coeff_sem <- map(mod_stab_list[!mask_drop_lm],
  get_intercept_variance) %>%
  enframe(name = "variable") %>%
  unnest(value) %>%
  pivot_wider(names_from = "name", values_from = "value")
get_intercept <- function (mod) {
  mod[["coefficients"]]$fixed["(Intercept)"]
}
intercept_sem <- map_dbl(mod_stab_list[!mask_drop_lm],
  get_intercept) %>%
enframe(name = "variable", value = "Intercept")

formated_random_effects <- intercept_sem %>% 
  left_join(random_coeff_sem, by = "variable") %>%
  rename(`Response variable` = variable)
#format_random <- function (x, y) {
  #tibble(
    #Variable = y, 
    #`Hydrographic basin` = rownames(x),
    #Intercept = x[[1]]
  #)
#}
#formated_random_effects <- do.call(rbind, map2(random_coeff_sem, names(random_coeff_sem), format_random))

caption_random_effect <- paste0(
  "Estimated random effect of hydrographic basin on the intercept of the linear
  models performed in structural equation modeling."
)

@@

### Robustness to different dataset selection

```{r sensitivity, message = FALSE, fig.dim = c(10, 7), fig.cap = "Sensivity analysis with different dataset", cache = FALSE}
library(furrr)

@@

sem_stab_l_data <- 
  furrr::future_map(sem_sensivity_data,

  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_stab_sem_rich
  )
  )

@@

```{r sem-sens-bm, cache = TRUE, fig.cap = "Sensivity analysis for Biomass"}
sem_bm_l_data <- furrr::future_map(sem_sensivity_data,
  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_prod_sem_rich)
)
sem_bm_results <- map(sem_bm_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

@@

```{r, fig.height = 8, fig.cap = sens_cap}
p_sens_bm <- plot_sensivity_analysis(.data = sem_bm_results) +
  theme(legend.position = "none")

p_sens_stab <- plot_sensivity_analysis(.data = sem_stab_results) +
  guides(color = guide_legend(direction = "horizontal", nrow = 2))
legend_b <- get_legend(p_sens_stab)
p_sens_stab %<>% + theme(legend.position = "none")
p <- plot_grid(
  p_sens_bm, p_sens_stab, legend_b, nrow = 3,
  rel_heights = c(1, 1, .1),
  labels = c("Total biomass", "Biomass stability")
)
p
```

>>>>

---
title: Appendix of Species richness and network structure jointly drive total biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault and Colin Fontaine
date: \today 
output:
  pdf_document:
    fig_caption: true 
    keep_tex: true
    toc: true
fontsize: 12pt
header-includes:
   -  \newcommand{\beginsupplement}{\setcounter{table}{2}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} 
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{float}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
---

@@

library(tidyverse)
library(magrittr)
library(ggeffects)
library(cowplot)
library(kableExtra)
library(sf)
library(nlme)
source(mypath("R", "misc.R"))
source(mypath("R", "plot_methods.R"))
source(mypath("R", "community_analysis.R"))
source(mypath("R", "community_methods.R"))
source(mypath("R", "statistical_analysis.R"))
source(mypath("R", "total_sem_effect.R"))
source(mypath("R", "press_methods.R"))
source(mypath("R", "geo_methods.R"))
source(mypath("R", "synchrony.R"))
source(mypath("R", "cleaning_methods.R"))

@@

\beginsupplement

@@

```{=latex}
\pagebreak
```

@@

Diet information were obtained from fishbase.org and literature (Appendix A, Table S2). The
diet information contained from two to three life stages by species, the life
stages being defined by the body size of the fishes. Over the 57 fish species,
only seven had a piscivorous life stage and only three had as strict piscivorous
life stage. Fish species were largely omnivorous.

```{r food, eval = FALSE}
onto_shift_cap <- paste0(
  "Ontogenic diet shift for each fish species.", 
  "Species had between two and three life stages, bounded by body size (in mm).
  A life stage of a given determine its diet. The fourth column to the last one
  correspond to this respective resource: light, detritus, biofilm,
  phytobenthos, macrophages, phytoplankton, zooplankton,  zoobenthos and
  fishes."
)
table_onto_diet <- fish_diet_shift %>%
  rename(`Life stage` = stage) %>%
  mutate(
    min = paste0("[", size_min),
    max = paste0(size_max, "]")
  ) %>%
  unite("Size interval", min:max, sep = ";") %>%
  select(species, `Life stage`, `Size interval`, light:ref_ods)

kable(table_onto_diet,
  format = "latex",
  booktabs = T,
  label = paste0("onto-diet"),
  caption = onto_shift_cap
  ) %>%
kable_styling(latex_options =c("striped", "scale_down"))
```

```{=latex}
%\includepdf[pages={2-7}]{figs/supplementary_ref_willem.pdf}
```

@@

```{r, fig.height = 20/2.54, fig.cap = pressure_cap}
pca_pressure <- habitat_press
myload(fourier, dir = mypath("data-raw", "polluants"))
myload(flow_temp_metrics, station_analysis, the_8_hydrologic_basin, dir = mypath("data"))

@@

map_pressure <- map(var_to_plot,
~p +
  geom_sf(data = sf_pressure, aes_string(color = .x), size = .5) +
  viridis::scale_color_viridis() +
  labs(title = str_replace_all(.x, var_replacement)) +
  theme(
    legend.title = element_blank(),
    text = element_text(size = 8),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  ) +
 guides(color = guide_colourbar(barwidth = 0.5, barheight = 4))
)

@@

We reconstructed the stream network and the connection between sites with the
`openSTARS` R package [@kattwinkel_preparing_2020]. The reconstruction was done with a raster of digital
elevation model at a resolution of 250 meters [(IGN)](https://wxs.ign.fr/jvam1hsjm11u8voorw81v2xb/telechargement/prepackage/BDALTI-75M_PACK_FXX_2018-01-24%24BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15/file/BDALTIV2_2-0_75M_ASC_LAMB93-IGN69_FRANCE_2018-01-15.7z)
and a theoretical hydrologic network [@pella_reseau_2012].

@@

Then we modeled each variable as dependant only on the covariation between sites
and taking in account stream network. To model the autocovariance between sites,
we used the linear-sill and mariah functions for taildown and tailup
autocovariance, according to their prediction performance. The interpolations
were done with the `SSN` R package 
[@isaak_applications_2014, @hoef_ssn:_2014].

@@

```{=latex}
\pagebreak
```

# Data analysis

@@

```{r eig-plot}
rotated_eigen_values <- 
  tibble(
    name = paste0("RC", seq(1,14)),
    type = c(rep("Kept", 2), rep("No kept", 14 - 2)),
    value = pca_rotated$rotated$values
  )
eigen_rotated_plot <-
  ggplot(rotated_eigen_values,
    aes(x = reorder(name, -value) , y = value)) +
  geom_col(aes(fill = type)) +
  scale_fill_discrete(name = "Status") +
  labs(x = "Rotated component", y = "Eigen values") +
  theme(legend.position = "none",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank())
```

@@

```{r, eval=FALSE}
corr_pca_caption <- paste0(
  "Spearman correlation between environmental variables used in the PCA.
  The variables have been scaled before to compute the correlation."
)

@@

```{r rc2-map, eval = FALSE, fig.cap = caption_rc2_map}
p +
  geom_sf(data = rc2_sf, aes(color = log(value + abs(min(value)) + 1 ))) +
  viridis::scale_color_viridis() +
  labs(title = "RC2 (Log)") +
  theme(legend.title = element_blank())
```

@@

```{=latex}
\pagebreak
```

# Statistical analysis

```{r}
sem_variable_table <- tibble(
  variable = c(paste0("PCA axis", seq(1, 2)), "Species richness", "Connectance", "Average trophic level", "Synchrony", "CVsp", "Total biomass"),
  unit = c(rep("unitless", 2), "# by square meter", rep("unitless", 4), "gram by square meter"
    ),
  transformation = c(rep("Log base 10", 2), "Log base 10", rep("No", 2), rep("Log base 10", 3)),
  interpretation = c("linked to stream size",
    "linked to average temperature",
    "Total species richness over the temporal series",
    "Median of annual link density", "Median of annual average trophic level", "", "",
    "Median of annual total biomass")) %>%
kable(
  format = "latex",
  booktabs = T,
  caption = "Variable description and transformation"
  ) %>%
kable_styling(latex_options = c("striped"))
sem_variable_table
```

@@

---
# Species richness to community attributes 
---

To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. In particular, species richness decreased connectance
following results from @thebault_stability_2010; @winemiller_must_1989. But
theoretically, some ecological networks can show the inverse relationship
[@winemiller_must_1989]. (6) We hypothesized also that species richness will
modulate average weighted trophic level because of food web assembly processes.
A community with more fish species can support higher trophic level. Finally, we
set up causal relationships from community attributes to stability components.

@@

@@

@@

@@

kable(format_coef_table(stab_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_stab_sem_caption
  ) %>%
collapse_rows(columns = 1, valign = "top") %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

@@

kable(format_coef_table(bm_sem_rich),
  format = "latex",
  booktabs = T,
  caption = est_bm_sem_caption
  ) %>%
collapse_rows(columns = 1, valign = "top") %>%
kable_styling(latex_options = c("striped", "scale_down"))
```

```{r rsq-model}
rm_n <- str_remove_all(get_sem_var_name_replacement(), pattern = "\n")
names(rm_n) <- names(get_sem_var_name_replacement()) 
rsq_table <- map_dfr(mod_stab_list, piecewiseSEM::rsquared, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2)) %>%
  mutate(Response = str_replace_all(Response, rm_n)) %>%
  dplyr::select(Response, Marginal, Conditional, R.squared)
rsq_table_caption <- paste0(
  "R squared of the linear models used in the structural equation modeling."
)
```

@@

# get random effect 
get_intercept_variance <- function (mod) {
  vc <- VarCorr(mod)
  suppressWarnings(storage.mode(vc) <- "numeric")
  enframe(vc[1,])
}
random_coeff_sem <- map(mod_stab_list[!mask_drop_lm],
  get_intercept_variance) %>%
  enframe(name = "variable") %>%
  unnest(value) %>%
  pivot_wider(names_from = "name", values_from = "value")
#check
#all.equal(random_coeff_sem[[2]],(random_coeff_sem[[3]])^2)
#good

# get intercept 
get_intercept <- function (mod) {
  mod[["coefficients"]]$fixed["(Intercept)"]
}
intercept_sem <- map_dbl(
  mod_stab_list[!mask_drop_lm], get_intercept) %>%
  enframe(name = "variable", value = "Intercept")

formated_random_effects <- intercept_sem %>% 
  left_join(random_coeff_sem, by = "variable") %>%
  rename(`Response variable` = variable)
caption_random_effect <- paste0(
  "Estimated variance and standard deviation on the intercept due to the hydrographic basins."
)

@@

```{r sensitivity, message = FALSE, fig.dim = c(10, 7), fig.cap = "Sensivity analysis with different dataset"}
library(furrr)

@@

sem_stab_l_data <- 
  furrr::future_map(sem_sensivity_data,
    ~build_dataset_get_sem_coefficient(
      .op = .x,
      sem_fun = compute_stab_sem_rich
    )
  )

@@

```{r sem-sens-bm, fig.cap = "Sensivity analysis for Biomass"}
sem_bm_l_data <- furrr::future_map(sem_sensivity_data,
  ~build_dataset_get_sem_coefficient(
    .op = .x,
    sem_fun = compute_prod_sem_rich)
)
sem_bm_results <- map(sem_bm_l_data,
  ~.x$coefficients[-ncol(.x$coefficients)]) %>%
  enframe() %>%
  rename(dataset = name) %>%
  unnest(value)

@@

```{r, fig.height = 8, fig.cap = sens_cap}
p_sens_bm <- plot_sensivity_analysis(
  .data = sem_bm_results,
  inverse_label_order = TRUE) +
  theme(legend.position = "none", plot.margin = unit(c(0, 0, 0, 25), units = "mm"))

p_sens_stab <- plot_sensivity_analysis(.data = sem_stab_results,
  inverse_label_order = TRUE) +
  guides(color = guide_legend(direction = "horizontal", nrow = 2))
legend_b <- get_legend(p_sens_stab)
p_sens_stab %<>% + theme(legend.position = "none", plot.margin = unit(c(0,0,0,27), units = "mm"))
p <- plot_grid(
  p_sens_bm, p_sens_stab, legend_b, nrow = 3,
  rel_heights = c(1, 1, .1),
  labels = c("Total biomass", "Biomass stability")
)
p
```

```{=latex}
\pagebreak
```
